An Open Repository of Real-Time COVID-19 Indicators

### 2021-06-18 indicators/paper/paper.tex

\begin{enumerate}
\item \textit{Geo-wise correlations} (i.e., on a specific date, do values of the
  signal correlate with case rates across locations?): Formally, let $X_t$ and
  $Y_t$ be vectors of values of a particular signal and case rates, over all
  locations, on date $t$. The geo-wise correlation at time $t$ is defined as
  $\cor(X_t, Y_t)$. This examines whether a signal has the capability to help
  spot locations with high case rates at any given time.

\item \textit{Time-wise correlations} (i.e., at a specific location, do values
  of the signal correlate with case rates across time?): Formally, let $X_\ell$
  and $Y_\ell$ be vectors of a values of a particular signal and case rates,
  over all times, at location $\ell$. The time-wise correlation at location
  $\ell$ is defined as $\cor(X_\ell, Y_\ell)$. This examines whether changes in
  a signal over time correspond to changes in reported cases at the same
  location.
\end{enumerate}

@@

Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
by the five signals and COVID case rates (from JHU CSSE, smoothed using a 7-day
trailing average), from April 15, 2020 to April 15, 2021. This calculation is
performed over counties with at least 500 cumulative cases by the end of this
period, and at which all indicators are available (956 counties in total). The
large positive correlations suggest that these signals could be useful in
hotspot detection (identifying counties with relatively high COVID activity, at
a given time). Somewhat surprisingly, the survey-based CLI-in-community signal
shows the strongest correlations for much of the time period. This clearly
demonstrates the value of a large-scale survey such as CTIS for tracking
symptoms and case trends, especially when other data is unavailable.

Figure~\ref{fig:time_wise_correlation} summarizes time-wise correlations from
these five signals over the same time period, and for the same set of counties.
For each signal, we display the set of correlations that it achieves in
histogram form (more precisely, using a kernel density estimate). All signals
produce positive correlations in the majority of counties considered (with very
little mass in each estimated density being to the left of zero). The largest
correlations, in bulk, are achieved by the CHNG-COVID signal; the
CTIS-CLI-in-community is a close second, and the CHNG-CLI signal is third. There
are two noteworthy points:

@@

There are a number of open questions, and challenges that remain. Several
signals are subject to biases, such as survey sampling and nonresponse biases,
geographic differences in market share for medical claims data, or biases in the
population represented in mobile phone mobility data \cite{Coston:2021}. Claims
data tends also to be subject to biases during major holidays and other events
that change health-seeking behavior. Characterizing these biases will be
important for future research and operational systems that use these signals.
Several data sources are also subject to extensive revision and backfill, which
must be studied and modeled to enable effective real-time use of these sources
in forecasting and nowcasting systems. The breadth and unique features of the
COVIDcast API will help facilitate this and other related work, which will be
vital to advancing pandemic modeling and preparedness.

>>>>

\begin{enumerate}
\item \textit{Geo-wise correlations} (i.e., on a specific date, do values of the
  signal correlate with case rates across locations?): Formally, let $X_t$ and
  $Y_t$ be vectors of values of a signal and case rates, over all locations, on
  date $t$. The geo-wise correlation at time $t$ is defined as $\cor(X_t, 
  Y_t)$ (where here and throughout $\cor(\cdot,\cdot)$ denotes Spearman 
  correlation). This examines whether a signal has the capability to help spot
  locations with high case rates at any given time.

\item \textit{Time-wise correlations} (i.e., at a specific location, do values
  of the signal correlate with case rates across time?): Formally, let $X_\ell$
  and $Y_\ell$ be vectors of values of a signal and case rates, over all
  times, at location $\ell$. The time-wise correlation at location $\ell$ is
  defined as $\cor(X_\ell, Y_\ell)$. This examines whether changes in a signal
  over time correspond to changes in reported cases at the same 
  location.
\end{enumerate}

@@

Figure~\ref{fig:geo_wise_correlation} shows the geo-wise correlations achieved
by the five signals and COVID case rates (from JHU CSSE, smoothed using a 7-day
trailing average), from April 15, 2020 to April 15, 2021. This calculation is
performed over counties with at least 500 cumulative cases by the end of this
period, and at which all indicators are available (956 counties in total). The
large positive correlations suggest that these signals could be useful in
hotspot detection (identifying counties that have relatively high COVID
activity, at a given time). Somewhat surprisingly, the survey-based
CLI-in-community signal shows the strongest correlations for much of the time
period. This clearly demonstrates the value of a large-scale survey such as CTIS
for tracking symptoms and case trends, especially when other data is unavailable.

Figure~\ref{fig:time_wise_correlation} summarizes time-wise correlations from
these five signals over the same time period, and for the same set of counties.
For each signal, we display the set of correlations that it achieves in
histogram form (more precisely, using a kernel density estimate). All signals
produce positive correlations in the majority of counties considered (with very
little mass in each estimated density being to the left of zero). The largest
correlations, in bulk, are achieved by the CHNG-COVID signal; the
CTIS-CLI-in-community signal is a close second, and the CHNG-CLI signal is
third. There are two noteworthy points:

@@

There are a number of open questions, and challenges that remain. Several
signals are subject to biases, such as survey sampling and nonresponse biases,
geographic differences in market share for medical claims data, or biases in the
population represented in mobile device mobility data.
% \cite{Coston:2021}. RJT: commenting out because if we cite only this, then I 
% feel like we will have to provided references for the other sources of bias.
% Since otherwise it might look like we don't think there's credible work on the
% others. 
Claims data tends also to be subject to biases during major national holidays
and other events that change health-seeking behavior. Characterizing these
biases will be important for future research and operational systems that use
these signals.  Several data sources are also subject to extensive revision and
backfill, which must be studied and modeled to enable effective real-time use of
these sources in forecasting and nowcasting systems. The breadth and unique
features of the COVIDcast API will help facilitate this and other related work,
which will be vital to advancing pandemic modeling and preparedness.
